---
title: "Quick Start: Metadata"
author: "Michael Koohafkan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start 2: Metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document provides a brief demonstration of how to query monitoring
program metadata with `wqpr`. The examples in this document will use
the MARSH program data on the production database:

```{r}
wqp_use_database("production")
wqp_use_program("marsh")
```

Metadata covered in this document includes basic station information
and contacts (staff with access to WQP). This information is used by
multiple features in the database.

The function `wqp_stations()` lists the stations available in WQP:

```{r}
wqp_stations()
```

The function `wqp_locations()` lists station "locations", i.e., places
where equipment can be installed at a station.

```{r}
wqp_locations()
```

The function `wqp_contacts()` lists staff associated with WQP.

```{r}
wqp_contacts()
```
