---
title: "Quick Start: Results"
author: "Michael Koohafkan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start 3: Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document provides a brief demonstration of how to query result data
(continuous or discrete) with `wqpr`. Usage of `wqpr` functions are
enhanced by the `tidyverse` (most notably the `dplyr` and `tidyr` packages)
and date/time parsing utilities provided by `lubridate`.

```{r message = FALSE}
library(tidyverse)
library(lubridate)
```

## Description

A "result" dataset is a collection of time-stamped numeric values
(and quality control flags) at a specific location. This could be
a continuous time series of of water quality or data collected
intermittently or periodically at a specific location. A result
dataset has the following properties:
- station
- analyte
- constituent (combination of analyte, equipment, and measurement units)
- reading type (e.g., time series, visit data, etc.)
- measurement interval (e.g., 15 minute)
- aggregation method (e.g., instantaneous, average, minimum, etc.)
 - start and end date

All result datasets are identified by an integer code (a "result id").
The result id is unique, but not informative; generally the user will
have knowledge of other aspects of the data (the station, analyte, 
etc.) which are used to identify the result data of interest. A 
complete listing of result ids and associated information can be
retrieved using the function `wqp_result_details()`.

## Querying result data

In this example, we will retrieve continuous (15-minute) 
salinity data from National Steel. We know the following
information about the result ID of interest:

 - station name: National Steel
 - cdec code: NSL
 - analyte: Specific Conductance
 - reading type: Time Series
 - interval: 15 min

The National Steel water quality station is managed by the MARSH
program. Because we are only querying data, we will use the
production database:

```{r}
wqp_use_database("production")
wqp_use_program("marsh")
```

```{r include = FALSE}
conductance.ids = wqp_result_details() %>%
  filter(
    cdec_code == "NSL",
    analyte_name == "Specific Conductance",
    reading_type_name == "Time Series",
    interval_name == "15 min"
  )
```

We use the `filter()` function from `dplyr` to trim the result
details information to result IDs that match this criteria.
In this example, `r nrow(conductance.ids)` distinct result ids are
returned (some columns omitted from the table below for clarity):

```{r eval = FALSE}
conductance.ids = wqp_result_details() %>%
  filter(
    cdec_code == "NSL",
    analyte_name == "Specific Conductance",
    reading_type_name == "Time Series",
    interval_name == "15 min"
  )
conductance.ids
```

```{r echo = FALSE, results = "asis"}
conductance.ids %>%
  select(result_id, cdec_code, analyte_name, unit_name,
    start_date, end_date) %>%
  kable()
```

The function `wqp_result_data()` is used to retrieve the actual
datasets associated with a vector of result ids. We can also specify
start and end timestamps to pull a subset of the data; if a result id
does not contain any data matching that interval a zero-row table is
returned. By default the function returns a list of dataframes, but
the argument `bind = TRUE` can be used to automatically concatenate
the datasets. In this case, we choose to retrieve data for June 
through September of 2019:

```{r, eval = FALSE}
conductance.data = wqp_result_data(conductance.ids$result_id,
  start.date = "2019-06-01", end.date = "2019-09-30 23:59:59",
  bind = TRUE)
head(conductance.data, 10)
```

```{r, echo = FALSE, results = "asis"}
conductance.data = wqp_result_data(conductance.ids$result_id, 
  start.date = "2019-06-01", end.date = "2019-09-30 23:59:59",
  bind = TRUE)
kable(head(conductance.data, 10))
```

## Inserting result data

Two functions are provide for inserting data.
`wqp_update_result_flags()` provides a safe way to overwrite quality
control flags without the risk of modifying the actual data values.
`wqp_insert_result_data()` provides a more general framework for
adding or overwriting data values as well as the accompanying quality
control flags.

In this example, we will update the flags associated with Chlorophyll a
data from the Collinsville water quality station (CDEC code CSE) on
February 15th. We use the Test database to avoid modifying the true
qualified data. Inserting data to WQP requires an authorization token.
For more information about tokens, see the
[Setup Quickstart](quickstart-setup.html) vignette. 

```{r, include = FALSE}
wqp_use_database("test")
```

```{r, eval = FALSE}
wqp_use_database("test")
wqp_use_token("<your token>")
```

First, query the data. For simplicity, we'll pull only the first hour
of the day.

```{r}
chlorophyll.ids = wqp_result_details() %>%
  filter(
    cdec_code == "CSE",
    analyte_name == "Chlorophyll",
    reading_type_name == "Time Series"
  )

chlorophyll.data = wqp_result_data(chlorophyll.ids$result_id,
  start.date = "2019-02-15", end.date = "2019-02-15 01:00:00",
  bind = TRUE) 
```

Next, overwrite the quality control flag with a new value. We'll assume
the data has been reviewed and classified as "Good".

```{r}
chlorophyll.update = chlorophyll.data %>%
  mutate(qaqc_flag_id = "G") 
```

Finally, use `wqp_update_result_flags()` to update the flags. We use
the `overwrite` argument to only update the flags of data points that
are currently marked as "unchecked".

```{r}
wqp_update_result_flags(chlorophyll.update, overwrite = "U")
```

If we pull the data again, we can confirm that the flags have been
updated.

```{r}
wqp_result_data(chlorophyll.ids$result_id,
  start.date = "2019-02-15", end.date = "2019-02-15 01:00:00",
  bind = TRUE) 
```

Inserting data values is accomplished in similarly using
`wqp_insert_data()`. In this example we overwrite existing
data, but the same approach can be used to insert new data.
The `overwrite` argument can be used to protect against accidental
overwrites, e.g., setting `overwrite = FALSE` allows new values
to be added but will not allow existing data to overwritten.

```{r}
chlorophyll.overwrite = chlorophyll.data %>%
  mutate(value = -10, qaqc_flag_id = "B")
wqp_insert_result_data(chlorophyll.overwrite, overwrite = TRUE)
```

If we pull the data again, we can confirm that the flags have been
updated.

```{r}
wqp_result_data(chlorophyll.ids$result_id,
  start.date = "2019-02-15", end.date = "2019-02-15 01:00:00",
  bind = TRUE) 
```

```{r include = FALSE}
# undo the previous examples
wqp_insert_result_data(chlorophyll.data, overwrite = TRUE)
```
